### 2022-07-15 scripts/windrose.Rmd
add windrose md

---
title: "Wind Rose"
author: "Marion Nyberg"
date: '2022-07-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(data.table)
library(janitor)
library(plotly)
library(plyr)
library(dplyr)
library(knitr)
library(tidyverse)
#library(devtools)
library(openair)
library(tidyr)
library(naniar)
library(forecast)
library(kableExtra)
library(htmltools)
library(ggpubr)
library(rstatix)
library(broom)
library(gtable)
library(gridExtra)
library(cowplot)
library(viridis)
library(zoo)
library(clifro)
library(lattice)

rm(list = ls())
#wd <- ("R:/SET/PlantSci/Shared Work Spaces/Ecology/Flux/SP_data")
#setwd(wd)
#knitr::opts_knit$set(root.dir = wd)
```

Below are conversion factors and functions used in the markdown file:

```{r functions and conversions}
d.avg <- 1800 * 48 #60 secs * 60 mins * 24 hours (# of secs in day)
co2_conv <- 12.01/(10^6)

tidy.csv <- function(x){
  input <- x[-c(1,2),] %>%
    row_to_names(., 1)
}

gofC.f <- function(x){ # function to convert to mean 30 minute to g of C
  (x*d.avg)*co2_conv
}

gtomg.f <- function(x){ # function to convert from g to mg
  x*1000
}
```

```{r read data and some data class cleaning, include=FALSE}
input <- fread("R:/SET/PlantSci/Shared Work Spaces/Ecology/Flux/SP_data/PFP/L6/SilverPlains_L6.csv", header = FALSE, na.strings = "-9999") 
input <- tidy.csv(input)

input <- input %>% select(xlDateTime, Fe, Fg_Av, Fh, Fld, Flu, Fn, Fsu, Fsd, Precip, RH, Sws, Ta, Ts, VPD, Wd_SONIC_Av, Ws_SONIC_Av,ER_LT_all, GPP_LT,GPP_SOLO, NEE_LT, ET) # select met variables (add more as needed)
colnames(input) <- c('Timestamp', 'LE', 'G', 'H', 'LWin', 'LWout', 'NETRAD', 'SWout', 'SWin', 'Precip', 'RH', 'SWC', 'Ta', 'Ts', 'VPD', 'WD', 'WS', 'ER_LT', 'GPP_LT', 'GPP_SOLO', 'NEE_LT', 'ET')

input$Timestamp <- as.POSIXct(input$Timestamp, format= "%d/%m/%Y %H:%M") # change from character to POSIXct
input[,2:22] <- as.data.frame(sapply(input[,2:22], as.numeric)) # convert data frame from character to numeric

input$jday <-yday(input$Timestamp) # add jday
input$Year <- year(input$Timestamp) # extract year
input$Year <- as.factor(input$Year) # convert year from integer to factor
input$month <- month(as.POSIXct(input$Timestamp)) # extract month
input$month <- as.factor(input$month)
input$hour <- format(input$Timestamp, format = "%H")
input$hour <- as.numeric(input$hour)

# create index for time of day - morn, aft, night
input$TOD <- ""
ind.morn<- (input$hour >= 06 & input$hour < 12)
ind.aft<- (input$hour >= 12 & input$hour < 18)
ind.night<- (input$hour >= 18 | input$hour < 6)

input$TOD[ind.morn] <- "morn"
input$TOD[ind.aft] <- "aft"
input$TOD[ind.night] <- "night"

input <- input[-c(13835, 13836, 31307, 31308)]
```

```{r add seasons, include=FALSE}
# maybe this would be better as GS/NGS?

input$season <- "" # add season variable

# define seasons
winter <- (input$month == "6" | input$month == "7" | input$month == "8")
spring <- (input$month == "9" | input$month == "10" | input$month == "11")
summer <- (input$month == "12" | input$month == "1" | input$month == "2")
autumn <- (input$month == "3" | input$month == "4" | input$month == "5")

input$season[winter] <- "Winter"
input$season[spring] <- "Spring"
input$season[summer] <- "Summer"
input$season[autumn] <- "Autumn"

season_levelorder <-  c('Summer', 'Autumn', 'Winter', 'Spring') 
```

### Wind direction and speed

```{r Wind direction, echo=FALSE, message = FALSE, fig.height=21, fig.width=12}
input$Month.WR <- input$month
#plot_ly(input, x = ~jday, y = ~WD, split = ~Year, name = ~Year, type = 'scatter', mode = 'lines')
windRose(input, ws = "WS", wd = "WD", type = "Year")

input%>%filter(month == "1") %>%
  windRose(., ws = "WS", wd = "WD", type = "TOD")

Windrose <- windRose(input, ws = "WS", wd = "WD", type = c("TOD","Month.WR"), aspect = "fill")

Windrose
jpeg("Windrose.jpeg", units="in", width=7, height=23,  res=300)
print(Windrose)
dev.off()

input %>% filter(season == "Autumn") %>%
  windRose(., ws = "WS", wd = "WD")

input %>% filter(season == "Winter") %>%
  windRose(., ws = "WS", wd = "WD")

input%>% filter(season == "Spring") %>%
  windRose(., ws = "WS", wd = "WD")
```

```{r}
# alternative way: https://www.r-bloggers.com/2016/06/plot-a-wind-rose-in-r/

# choose bin size (degrees/bin)
deg <- 30 
# define the range of each bin
dir.breaks <- seq(0-(deg/2), 360+(deg/2), deg)
# assign each direction to a bin range
dir.binned <- cut(input$WD,
                       breaks = dir.breaks,
                       ordered_result = TRUE)
# generate pretty lables
dir.labels <- as.character(c(seq(0, 360-deg, by = deg), 0))
# replace ranges with pretty bin lables
levels(dir.binned) <- dir.labels
# Assign bin names to the original data set
input$dir.binned <- dir.binned

thm <- theme_bw() + 
    theme(axis.text.x = element_text(size=8, face = "plain"),
          axis.text.y = element_text(size=8, face = "plain"),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size=8, face = "plain", hjust = 0.9, vjust = 1.3),
          panel.border = element_blank(),
          panel.grid  = element_blank())

summary(dir.binned) # see which bin is most common

# initialise the plot
plt.dirrose <- ggplot() +
    # since the plot background is blank we'll add a series of horizontal lines, at 5000 count intervals, up to 25000.
    geom_hline(yintercept = seq(0, 150, by = 25), colour = "grey60", size = 0.3) +
    # Now we add a darker horizontal line as the top border at 30000.
    geom_hline(yintercept = 150, colour = "black", size = 0.3) +
    # We want 12 vertical lines representing the centers of the 30° ranges.
    geom_vline(xintercept = c(seq(1,12,1)), colour = "grey60", size = 0.3) +
    # On top of everything we place the histogram bars.
    geom_bar(data = input, aes(x = dir.binned), width = 1, colour="black", size = 0.3, alpha=0.5) + 
    # Add the x-axis labels
    scale_x_discrete( drop = FALSE, labels = c(0, "", "", 90, "", "", 180, "", "", 270, "", "")) +
    # Add the y-axis labels
    scale_y_continuous(limits = c(0, 150), expand = c(0, 0), 
                       breaks = c(0, 25, 50, 75, 100, 125, 150), 
                       labels = c(0, 25, 50, 75, 100, 125, 150)) +
    # Add the axis titles
    labs(x = 'Outward step bearing (°)', y = 'Count of outward steps (x10³)') +
    # If you only use the plot code up till here you will get a histogram.
    # the next line wraps the histogram into a windrose
    coord_polar(start = -(deg/2)*(pi/180)) +
    # apply theme
    thm
plt.dirrose
```

```{r wind and fluxes, fig.width= 10}
# calculate an average wind pattern as in Griebel, A., et al., 2016
# wind pattern (WP) is equivalent to a standardized wind sector contribution, based on some temporal average

# there is way tooo much code here and I need to tidy it a lot and also do by season as well as year.

# define wind sectors
NE <- input.daily$WD >0 & input.daily$WD <= 90
SE <- input.daily$WD >90 & input.daily$WD <= 180
SW <- input.daily$WD >180 & input.daily$WD <= 270
NW <- input.daily$WD >270 & input.daily$WD <= 360

# adding wind sectors to daily df
input.daily$wind.sector <- ""
input.daily$wind.sector <- ifelse(input.daily$WD >0 & input.daily$WD <= 90, "NE", input.daily$wind.sector)
input.daily$wind.sector <- ifelse(input.daily$WD >90 & input.daily$WD <= 180, "SE", input.daily$wind.sector)
input.daily$wind.sector <- ifelse(input.daily$WD >180 & input.daily$WD <= 270, "SW", input.daily$wind.sector)
input.daily$wind.sector <- ifelse(input.daily$WD >270 & input.daily$WD <= 360, "NW", input.daily$wind.sector)

# calculate number of occurrences (n) per wind sector (j)
# need to make function to do this
NE.n <- input.daily %>% filter(Year == 2020 | Year == 2021) %>% filter(wind.sector == "NE") %>% nrow()
NE.n.2021 <- input.daily %>% filter(wind.sector == "NE" & Year == "2021") %>% nrow()
NE.n.2020 <- input.daily %>% filter(wind.sector == "NE" & Year == "2020") %>% nrow()

SE.n <- input.daily %>% filter(Year == 2020 | Year == 2021) %>% filter(wind.sector == "SE") %>% nrow()
SE.n.2021 <- input.daily %>% filter(wind.sector == "SE" & Year == "2021") %>% nrow()
SE.n.2020 <- input.daily %>% filter(wind.sector == "SE" & Year == "2020") %>% nrow()

NW.n <- input.daily %>% filter(Year == 2020 | Year == 2021) %>% filter(wind.sector == "NW") %>% nrow()
NW.n.2021 <- input.daily %>% filter(wind.sector == "NW" & Year == "2021") %>% nrow()
NW.n.2020 <- input.daily %>% filter(wind.sector == "NW" & Year == "2020") %>% nrow()

SW.n <- input.daily %>% filter(Year == 2020 | Year == 2021) %>% filter(wind.sector == "SW") %>% nrow()
SW.n.2021 <- input.daily %>% filter(wind.sector == "SW" & Year == "2021") %>% nrow()
SW.n.2020 <- input.daily %>% filter(wind.sector == "SW" & Year == "2020") %>% nrow()

# number of occurrences per wind sector divided by number of years (N). N is 2.1 because I did number of days divided by 365 to get the number of years (maybe better to filter to where I have only 2 full years)
WP.NE <- NE.n/2.1
WP.SE <- SE.n/2.1
WP.NW <- NW.n/2.1
WP.SW <- SW.n/2.1

# average carbon flux from each sector during each year
NE.co2.2021 <- input.daily %>% filter(wind.sector == "NE" & Year == "2021") %>% select(Fco2) %>%
  summarise(sum(Fco2)) # 5 obs 
NW.co2.2021 <- input.daily %>% filter(wind.sector == "NW" & Year == "2021") %>% select(Fco2) %>%
  summarise(sum(Fco2)) # 132 obs summarise(sum(Fco2)) 
NE.co2.2020 <- input.daily %>% filter(wind.sector == "NE" & Year == "2020") %>% select(Fco2) %>%
  summarise(sum(Fco2)) # 4 obs summarise(sum(Fco2)) 
NW.co2.2020 <- input.daily %>% filter(wind.sector == "NW" & Year == "2020") %>% select(Fco2) %>%  #140 obs
summarise (sum(Fco2))

SE.co2.2021 <- input.daily %>% filter(wind.sector == "SE" & Year == "2021") %>% select(Fco2) %>%
  summarise(sum(Fco2)) 
SW.co2.2021 <- input.daily %>% filter(wind.sector == "SW" & Year == "2021") %>% select(Fco2) %>%
  summarise(sum(Fco2)) 
SE.co2.2020 <- input.daily %>% filter(wind.sector == "SE" & Year == "2020") %>% select(Fco2) %>%
  summarise(sum(Fco2)) 
SW.co2.2020 <- input.daily %>% filter(wind.sector == "SW" & Year == "2020") %>% select(Fco2) %>%  
summarise (sum(Fco2))

# need to rename direction variable

NEco2.2020 <- -0.48/4
NEco2.2021 <- -2.09/5
NWco2.2020 <- 46.34/140
NWco2.2021 <- 30.67/132

SEco2.2020 <- -14.4/57
SEco2.2021 <- -49.31/55
SWco2.2020 <- -23.22/162
SWco2.2021 <- -82.32/169

Avg.wind.co2 <- data.frame(NEco2.2020, NEco2.2021, NWco2.2020, NWco2.2021,SEco2.2020, SEco2.2021,SWco2.2020, SWco2.2021) %>% pivot_longer(cols = NEco2.2020:SWco2.2021, names_to = "Windsec", )
Avg.wind.co2$Year <- ""
Avg.wind.co2$Year <- c("2020", "2021", "2020", "2021", "2020", "2021","2020", "2021")

Avg.wind.co2 %>% ggplot(aes(Windsec, value, group = Year)) + geom_line(aes(group = Year, colour = Year))

# NW is on  average a CO2 source and SE largest CO2 sinks
```